---
title: "628 module 2"
subtitle: "STAT 628"
author: "Group 1"
date: "Oct. 18, 2021"
output:
 html_document:
   df_print: paged
   toc: true
   toc_depth: 3
   number_sections: false
   theme: default
   highlight: tango
   toc_float:
     collapsed: true
     smooth_scroll: true
geometry: left = 3.18cm, right = 3.18cm, top = 2.0cm, bottom = 2.0cm
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE, tidy=FALSE)
options(scipen=999)
```



```{r}
library(car)
library(ggm)
library(corpcor)
library(nlme)
library(ggplot2)
```

## Data cleaning

```{r}
data <- read.csv("BodyFat.csv")
summary(data)
```

Two observations have implausible body fat values: 172 and 182

```{r}
data[data[,"BODYFAT"] < 3,]
```

Five observations have incorrect density or body fat measurements based on Siri's equation: 48, 76, 96, 182, 216

```{r}
data$SIRIS_ERROR = (495/data$DENSITY)-450 - data$BODYFAT
data[abs(data[,"SIRIS_ERROR"]) > 2,]
```


BMI is weight/height^2 * 703

Three rows have incorrect BMI, weight, or height: 42, 163, 221

```{r}
data$BMI_ERROR = data$WEIGHT/((data$HEIGHT)^2)*703 - data$ADIPOSITY
data[abs(data[,"BMI_ERROR"]) > 2,]
```

Nine rows need to be replaced or imputated: 42, 48, 76, 96, 163, 172, 182, 216, 221
Since this is only 3.6% of the data and it seems to be inaccurate at random, we'll simply remove these rows. 


```{r}
newdata = data[-c(42, 48, 76, 96, 163, 172, 182, 216, 221), ]
newdata = subset(newdata, select = -c(SIRIS_ERROR,BMI_ERROR))
newdata
```

Writing the cleaned up data to a csv:

```{r}
write.csv(newdata, "bodyfat_cleaned.csv")
```

## Deal with data

```{r}
# read the data 
BodyFat_clean <- read.csv('bodyfat_cleaned.csv',sep = ',')
BodyFat_clean <- BodyFat_clean[,c(3,5:18)]
```



```{r}
# look the data
head(BodyFat_clean)
tail(BodyFat_clean)

summary(BodyFat_clean)
```


```{r}
# plot multi-scatterplot
scatterplotMatrix(BodyFat_clean)

# correlation
cor(BodyFat_clean)

# partial correlation
cor2pcor(cor(BodyFat_clean))

```


```{r}
# divide age into group varible
AGE.GROUP <- NULL

for (i in 1:length(BodyFat_clean$AGE)) {
  if (BodyFat_clean$AGE[i] < 30) {
    AGE.GROUP[i] <- 1
  }
  else if (BodyFat_clean$AGE[i] > 60){
    AGE.GROUP[i] <- 2
  }
  else{
    AGE.GROUP[i] <- 3
  }
}

AGE.GROUP <- as.factor(AGE.GROUP)

BodyFat_clean <- cbind(BodyFat_clean,AGE.GROUP)
```






## Rule of Thumb

1. With larger R2, the model is better
2. The number of parameters in the model is usually not greater than 5.

## Fit models

```{r}
# fit models
attach(BodyFat_clean)

#based on correlation
m1 <- lm(BODYFAT~ ABDOMEN)
summary(m1)

m2 <- lm(BODYFAT~ ABDOMEN + ADIPOSITY + CHEST)
summary(m2)

#based on partial correlation
m3 <- lm(BODYFAT~ AGE  + ADIPOSITY + ABDOMEN+ CHEST + WRIST)
summary(m3)

m4 <- lm(BODYFAT~ AGE  + log(ADIPOSITY) + log(ABDOMEN) + log(CHEST) + log(WRIST))
summary(m4)

m5 <- lm(BODYFAT~ AGE  + log(ABDOMEN) + log(WRIST))
summary(m5)

m6 <- lm(BODYFAT ~ AGE.GROUP +  log(ABDOMEN) + log(WRIST),data = BodyFat_clean)
summary(m6)
detach(BodyFat_clean)
```

Compare all the models, m6 has the large R2, and just have three variables.


## Visualizing

```{r}
# Visualizing multiple linear regression models

ggplot(subset(subset(BodyFat_clean,AGE.GROUP  == 1)),aes(y=BODYFAT,x=log(ABDOMEN),color=log(WRIST)))+geom_point()+stat_smooth(method="lm",se=TRUE)

ggplot(subset(subset(BodyFat_clean,AGE.GROUP  == 2)),aes(y=BODYFAT,x=log(ABDOMEN),color=log(WRIST)))+geom_point()+stat_smooth(method="lm",se=TRUE)

ggplot(subset(subset(BodyFat_clean,AGE.GROUP  == 3)),aes(y=BODYFAT,x=log(ABDOMEN),color=log(WRIST)))+geom_point()+stat_smooth(method="lm",se=TRUE)

```


## Robust

```{r}
## test the robust of model

# Divide the data
n <- nrow(BodyFat_clean)
mark <- rep(FALSE, n)
mark[sample(n, round(2*n/3))] <- TRUE
training.set <- BodyFat_clean[mark, ]
test.set <- BodyFat_clean[!mark, ]

m <- lm(BODYFAT ~ AGE.GROUP +  log(ABDOMEN) + log(WRIST),data = training.set)
summary(m)

# test the model
test <- predict(m,test.set)

R2 <- 1 - sum((test.set$BODYFAT - test)^2)/sum((test.set$BODYFAT - mean(test.set$BODYFAT))^2)
R2.adj <- 1 -  (1 - R2 ) * (nrow(test.set)-1)/(nrow(test.set) - length(test.set) - 1)

R2
R2.adj
```




## Diagnosis the data

```{r}

stu.residual <- rstudent(m6)

# normality
qqPlot(stu.residual , main="qq plot", col="blue", col.lines="red",xlab = "Theoretical Quantiles",ylab = "Student residual")

# equal variance and linearity
ggplot( mapping = aes(x = m6$fitted.values, y = stu.residual)) + geom_point() + geom_line(y = 0,col = "red") + labs(x = "fitted.values",y= "Student residual")

# outiler
cook <- cooks.distance(m6)
head(sort(cook,decreasing = TRUE)) # Rule of Thumb is to compare cook distance with 1

# independece
ggplot( mapping = aes(y = stu.residual,x = c(1:length(stu.residual)))) + geom_point()+ geom_line(y = 0,col = "red") + labs(x = "index",y= "Student residual") + geom_line()

# Multicollinearity
vif_i <- vif(m6)

maxvif <- max(vif_i[,3])
maxvif

vif <- mean(vif_i[,3])
vif
```

## Remedy the model

Thus, in the qqplot and residual plot, there are three points anormly. We remove them to remedy the model. 

The max of the largest VIF value among $VIF_i$ is not larger than 10, and the mean vif is not considerablygreater than 1, but nearly equal to 1, so there is not multicollinearity.


```{r}

new <- BodyFat_clean[-39,]

m <- lm(BODYFAT ~ AGE.GROUP +  log(ABDOMEN) + log(WRIST),data = new)

summary(m)

stu.residual <- rstudent(m)

# normality
qqPlot(stu.residual , main="qq plot", col="blue", col.lines="red",xlab = "Theoretical Quantiles",ylab = "Student residual")

# equal variance and linearity
ggplot( mapping = aes(x = m$fitted.values, y = stu.residual)) + geom_point() + geom_line(y = 0,col = "red") + labs(x = "fitted.values",y= "Student residual")

# outiler
cook <- cooks.distance(m)
head(sort(cook,decreasing = TRUE)) # Rule of Thumb is to compare cook distance with 1


```

Compare the initial model, the coefficient just change a little, and the plots also does not change much. We can conclude that those points are not influential. Thus, we do not need to change the model.





















